---
title: "Titanic Dataset"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
---

```{r setup, include=FALSE}
library(flexdashboard)
library(data.table)
library(ggplot2)
library(plotly)
library(tidyverse)
library(DT)
library(ISLR)

path<-getwd()

train_data<-fread(paste(path,"data","train.csv",sep="/"))
test_data<-fread(paste(path,"data","test.csv",sep="/"))


train_data[,':='(Survived_class=gsub(1,"Survived",Survived))][
  ,':='(Survived_class=gsub(0,"Died",Survived_class))]

train_melt<-melt(train_data,id.vars =c("PassengerId","Survived","Pclass","Name",
                                       "Sex","Ticket","Cabin","Embarked"))


live_colors<-c("Survived"="darkgreen","Died"="darkred")
sex_colors<-c("female"="#FF33F0","male"="#2244D0")


```



Dataset description
=====================================

Row
-----------------------------------------------------------------------

### Intro
The Titanic dataset is composed by two groups:

- Train dataset
- Test dataset

Row
-----------------------------------------------------------------------

### Number of Measurements
```{r}

merged<-rbind(train_data[,':='(dataset="Train data")],
              test_data[,':='(dataset="Test data")],fill=TRUE)
data_desc<-merged[,.N,by=dataset]%>%
  ggplot()+
  geom_col(aes(x=dataset,y=N),fill="darkblue")+
  theme_bw()+
  ggtitle("Passengers per Dataset (N)")+
  xlab("")+ylab("Number of Passengers")

ggplotly(data_desc)

```

### Variable Types
```{r}
datatable(t(train_data[,lapply(.SD, class)]))
```

Row
-----------------------------------------------------------------------


### Gender Pie
```{r}
sex_count<-train_data[,.N,by=Sex]

plot_ly(sex_count, labels = ~Sex, values = ~N, type = 'pie')%>%
  layout(title="Gender Distribution - train data")
```


### Class Pie
```{r}
sex_count<-train_data[,.N,by=Pclass]

plot_ly(sex_count, labels = ~Pclass, values = ~N, type = 'pie')%>%
  layout(title="Ticket Class Distribution - train data")
```


### Survived Pie
```{r}
sex_count<-train_data[,.N,by=Survived_class]

plot_ly(sex_count, labels = ~Survived_class, values = ~N, type = 'pie')%>%
  layout(title="Survival Distribution - train data")
```


Row
-----------------------------------------------------------------------

###
All numerical variables are going to be analysed taking into consideration the survived column. The categorical variables will be evaluated in terms of relevance for the percentage of survived people.



Univariate View
=====================================


Row
-----------------------------------------------------------------------

###
In this section only numeric variables are presented and its impact on the survived status is preliminary evaluated.


Row
-----------------------------------------------------------------------

### Gender effect on survival status - Absolute Counts
```{r}
gender<-train_data[,.N,by=.(Sex,Survived_class)]%>%
  ggplot()+
  geom_col(aes(x=Sex,y=N,fill=Survived_class,group=Survived_class))+
  theme_bw()+
  theme(legend.position = "top")+
  scale_fill_manual(values=live_colors)+
  ggtitle("Gender effect - Absolute Counts")
ggplotly(gender)%>%layout(legend = list(x = 0.1, y = 0.9))

```


### Gender effect on survival status - Percentage for total of people
```{r}
gender_prec<-train_data[,.N,by=.(Sex,Survived_class)][,':='(N=100*N/sum(N))]%>%
  ggplot()+
  geom_col(aes(x=Sex,y=N,fill=Survived_class,group=Survived_class))+
  theme_bw()+
  theme(legend.position = "top")+
  scale_fill_manual(values=live_colors)+
  ggtitle("Gender effect - Percentage")+
  ylab("N (%)")
ggplotly(gender_prec)%>%layout(legend = list(x = 0.1, y = 0.9))

```


Row
-----------------------------------------------------------------------

### Survived vs Class (ticket)
```{r}
plot3<-train_data[,.N,by=.(Survived_class,Pclass)]%>%
  ggplot()+
  geom_col(aes(x=Pclass,y=N,fill=Survived_class))+
  theme_bw()+
  theme(legend.position = "top")+
   scale_fill_manual(values=live_colors)

ggplotly(plot3)%>%layout(legend = list(x = 0.1, y = 0.9))


```



Row
-----------------------------------------------------------------------

### Survived vs Age
```{r}
 ggplotly(ggplot(train_data)+
  geom_histogram(aes(x=Age,fill=Survived_class),alpha=0.7)+
  theme_bw()+
  theme(legend.position = "top")+
    facet_grid(Sex~.)+
   scale_fill_manual(values=live_colors)+
    ggtitle("Age Distribution Sex and Survival \n")
    )%>%layout(legend=list(x=0.0,y=-0.2))


```



Row
-----------------------------------------------------------------------

### Survived vs Fare
```{r}
 ggplotly(ggplot(train_data,aes(x=Fare,fill=Survived_class))+
    geom_density(fill = 'Orange', alpha=0.4)+
  theme_bw()+
  theme(legend.position = "top")+
   scale_fill_manual(values=live_colors)+
    ggtitle("Fare Distribution")+
    facet_wrap(~Survived)+xlim(0,120)
    )%>%layout(legend=list(x=0.0,y=-0.2))


```


Logistic Regression
========================================

Row
--------------------------------------------

###
In this section a simple classification model is run in order to evaluate the variables that increases the most the probability of survival.


```{r include=FALSE}

train_data_clean<-train_data[,.(Survived,Pclass,Sex,Age,SibSp,Parch,Fare,Embarked)]
train_data_clean<-na.omit(train_data_clean)

logistic_fit <- glm(Survived ~ Pclass + Sex + Age + SibSp + Parch + 
                                            Fare + Embarked, 
                    data = train_data_clean,
                    family = binomial)

probab_vector <- predict(logistic_fit,type = "response")

train_data_clean[,':='(Probability=probab_vector)][,                                          ':='(Predicted=ifelse(Probability>0.6,"Died","Survived"))][
  ,':='(ROC_prep=ifelse(Predicted=="Died" & Survived=="0","TN",""))][,':='(ROC_prep=ifelse(Predicted=="Survived" & Survived=="1","TP",ROC_prep))]


```


Row
------------------------------------------------

###
```{r}
ggplotly(ggplot(train_data_clean,aes(x=Probability))+
  geom_histogram()+
  facet_wrap(~Survived)+
    theme_bw()+
    ggtitle("odds distribution for classification model"))
```



